---
title: "Simulating_Transects"
author: "Michelle DePrenger-Levin"
date: "2026-03-25"
output: html_document
---


```{r}

library(dsims)
library(mgcv)

```

Simulate based on field data  
```{r}
picnic.region <- make.region(region.name = "Picnic", shape = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScleroDistanceSampling/Picnic/Picnic.shp",
                            units = "m")
plot(picnic.region)

coverP <- make.coverage(picnic.region, n.grid.points = 1000)

```

```{r}

# density surface in relative terms. maintains relative density when given absolute values later 
# Test how much bias will be introduced by non-uniform coverage
pdens <- make.density(picnic.region, 
                      x.space = 1, 
                      constant = 1) # relative density

pdens <- add.hotspot(object = pdens,
                     centre = c(732485, 4283590),
                     sigma = 60,
                     amplitude = 4)

pdens <- add.hotspot(object = pdens,
                     centre = c(732400, 4283545),
                     sigma = 10,
                     amplitude = 2)

pdens <- add.hotspot(object = pdens,
                     centre = c(732420, 4283575),
                     sigma = 10,
                     amplitude = 2)

pdens <- add.hotspot(object = pdens,
                     centre = c(732370, 4283560),
                     sigma = 30,
                     amplitude = -1)


plot(pdens)
pdensity.data <- pdens@density.surface[[1]]

## s() is a smooth term
fit.pgam <- gam(density ~ s(x,y), data = pdensity.data, family = gaussian(link="log"))
pgam.density <- make.density(region = picnic.region,
                            x.space = 1,
                            fitted.model = fit.pgam)

plot(pgam.density)
```


Population parameters    
Size as zero-truncated Poisson distribution with mean equal to 3   
Population size: N refers to the number of clusters rather than individuals. 
fixed.N = TRUE to generate population based on population size rather than density surface     
Trackline length is sum of lengths of transects plus off-effort transit time to complete the survey from beginning of first to end of last   
<https://cloud.r-project.org/web/packages/dssd/vignettes/GettingStarted.html> 
```{r}
## 1247 individuals at Picnic

trnc <- 3
scaleparam <- 1.16
shapeparam <- 2.58 
N <- trunc(1247) # listed as size of cluster but assume talks about size of population
reps = 999
# spacing <- 20# 25# 15# 20
desnAn <- 55

# covariates <- list(size = list(distribution = "ztruncpois", mean = 2)) ## Cactus can occur in groups but often as 1  
# Define the population description
pop.desc <- make.population.description(region = picnic.region,
                                        density = pgam.density,
                                        covariates = list(),
                                        N = N,
                                        fixed.N = TRUE)

# Covariate parameters on the log scale - what does this refer to???
# cov.param <- list(size = log(1.08))
detect <- make.detectability(key.function = "hr",
                             scale.param = scaleparam,   # 1.16
                             shape.param = shapeparam,   # 2.58
                             # cov.param = cov.param,
                             truncation = trnc)
plot(detect, pop.desc)


## hazard rate with no covariates, hazard rate with cluster size as a covariate
# with default error variance estimator R2: random line placement, unequal line lengths
#                                       R3: random, model-assisted estimator, true contagion process
#                                       S2: systematic, variances weighted by line length
analyses <- make.ds.analysis(dfmodel = list(~1),# list(~1, ~size),
                             key = c("hr"),  # ,"hr"),
                             truncation = trnc,
                             er.var = "R2",  # "S2": not supported, likely will be over-estimated, O2 also not supported
                             criteria = "AIC")
##############################################################################################################################
```


```{r}

summarySpacingOrintationType <- do.call(rbind,lapply(c(10,15,20,25), function(spacing){ 


##############################################################################################################################
## Parallel to gradient
picnic.design <- make.design(region = picnic.region,
                             design = "systematic",
                             # line.length = ,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn,
                             truncation = trnc,
                             spacing = spacing,
                             coverage.grid = coverP)


## Generate one random set of transects
transects <- generate.transects(picnic.design)

## Assess coverage uniformity and design stats   
# Want uniform coverage across the majority of the survey region
design <- run.coverage(picnic.design, reps = 1000)

save(design, file = paste0("Design",spacing,".Rdata"))

sim.parallel <- make.simulation(reps = reps,
                                design = picnic.design,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel <- run.survey(sim.parallel)

save(eg.parallel,  file = paste0("egSurveyspacing",spacing,".Rdata"))


sim.parallel <- run.simulation(sim.parallel, run.parallel = TRUE)
summarySim <- summary(sim.parallel, use.max.reps = TRUE)
est.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel@results$individuals$N)[3],function(x){
  out <- sim.parallel@results$individuals$N[,,x]
  out
}))  )

est.df <- est.df %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

save(sim.parallel, file = paste0(Sys.Date(),"PicnicParallelspacing",spacing,".Rdata"))

parallel.df <- data.frame(parORseg = "systematic",              # systematic grid or segments  
                     perpORparallel = "parallel",   # To the gradient
                     spacing = spacing,
                     designAngle = desnAn,
                     popsz = N,
                     numSamplers = transects@samp.count,
                     coverageAvg = transects@cov.area/transects@strata.area,
                     summarySim@individuals$summary,
                     summarySim@individuals$N,
                     summarySim@individuals$D,
                     est.df)

#######################################################################################################################################
##################################################
# Perpendicular to gradient, paralell to gradient contours
picnic.design.perp <- make.design(region = picnic.region,
                             design = "systematic",
                             # line.length = ,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn+90,
                             truncation = trnc,
                             spacing = spacing,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.perp <- generate.transects(picnic.design.perp)

## Assess coverage uniformity and design stats   
# Want uniform coverage across the majority of the survey region
design.perp <- run.coverage(picnic.design.perp, reps = 1000)

save(design.perp,  file = paste0("DesignPerp",spacing,".Rdata"))

sim.parallel.perp <- make.simulation(reps = reps,
                                design = picnic.design.perp,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel.perp <- run.survey(sim.parallel.perp)

save(eg.parallel.perp,  file = paste0("egSurveyPerpspacing",spacing,".Rdata"))

sim.parallel.perp <- run.simulation(sim.parallel.perp, run.parallel = TRUE)
summarySimPerp <- summary(sim.parallel.perp, use.max.reps = TRUE)
estperp.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel.perp@results$individuals$N)[3],function(x){
  out <- sim.parallel.perp@results$individuals$N[,,x]
  out
}))  )

estperp.df <- estperp.df %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

save(sim.parallel.perp, file = paste0(Sys.Date(),"PicnicPerpspacing",spacing,".Rdata"))

perp.df <- data.frame(parORseg = "systematic",              # parallel grid or segments  
                     perpORparallel = "perpendicular",   # To the gradient
                     spacing = spacing,
                     designAngle = desnAn+90,
                     popsz = N,
                     numSamplers = transects.perp@samp.count,
                     coverageAvg = transects.perp@cov.area/transects.perp@strata.area,
                     summarySimPerp@individuals$summary,
                     summarySimPerp@individuals$N,
                     summarySimPerp@individuals$D,
                     estperp.df )

##########################################################################
## Segments
## Parallel to gradient
picnic.design.seg <- make.design(region = picnic.region,
                             design = "segmentedgrid",
                             # line.length = ,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn,
                             truncation = trnc,
                             spacing = spacing,
                             seg.length = 30,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.seg <- generate.transects(picnic.design.seg)

## Assess coverage uniformity and design stats   
# Want uniform coverage across the majority of the survey region
design.seg <- run.coverage(picnic.design.seg, reps = 1000)

save(design.seg,  file = paste0("DesignSeg",spacing,".Rdata"))

sim.parallel.seg <- make.simulation(reps = reps,
                                design = picnic.design.seg,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel.seg <- run.survey(sim.parallel.seg)
save(eg.parallel.seg,  file = paste0("egSurveySegpspacing",spacing,".Rdata"))


sim.parallel.seg <- run.simulation(sim.parallel.seg, run.parallel = TRUE)
summarySeg <- summary(sim.parallel.seg, use.max.reps = TRUE)
estseg.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel.seg@results$individuals$N)[3],function(x){
  out <- sim.parallel.seg@results$individuals$N[,,x]
  out
}))  )

estseg.df <- estseg.df %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

save(sim.parallel.seg, file = paste0(Sys.Date(),"PicnicSegspacing",spacing,".Rdata"))

seg.df <- data.frame(parORseg = "segments",              # parallel grid or segments  
                     perpORparallel = "parallel",   # To the gradient
                     spacing = spacing,
                     designAngle = desnAn,
                     popsz = N,
                     numSamplers = transects.seg@samp.count,
                     coverageAvg = transects.seg@cov.area/transects.seg@strata.area,
                     summarySeg@individuals$summary,
                     summarySeg@individuals$N,
                     summarySeg@individuals$D,
                     estseg.df)


####################################################
# Perpendicular to gradient
picnic.design.seg.perp <- make.design(region = picnic.region,
                             design = "segmentedgrid",
                             # line.length = ,  ## or spacing, line length sets spacing to get close to the total length of all samplers
                             edge.protocol = "minus",      # don't go outside study area
                             design.angle = desnAn+90,
                             truncation = trnc,
                             spacing = spacing,
                             seg.length = 30,
                             coverage.grid = coverP)


## Generate one random set of transects
transects.seg.perp <- generate.transects(picnic.design.seg.perp)

## Assess coverage uniformity and design stats   
# Want uniform coverage across the majority of the survey region
design.seg.perp <- run.coverage(picnic.design.seg.perp, reps = 1000)

save(design.seg.perp,  file = paste0("DesignSegPerp",spacing,".Rdata"))


sim.parallel.seg.perp <- make.simulation(reps = reps,
                                design = picnic.design.seg.perp,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
eg.parallel.seg.perp <- run.survey(sim.parallel.seg.perp)

save(eg.parallel.seg.perp,  file = paste0("egSurveySegPerpspacing",spacing,".Rdata"))

sim.parallel.seg.perp <- run.simulation(sim.parallel.seg.perp, run.parallel = TRUE)
summarySegPerp <- summary(sim.parallel.seg.perp, use.max.reps = TRUE)
estsegperp.df <- as.data.frame(do.call(rbind,lapply(1:dim(sim.parallel.seg.perp@results$individuals$N)[3],function(x){
  out <- sim.parallel.seg.perp@results$individuals$N[,,x]
  out
}))  )

estsegperp.df <- estsegperp.df %>%
  filter(!is.na(Estimate)) %>%
  summarise(meanEst = mean(Estimate),
         meanLCL = mean(lcl),
         meanUCL = mean(ucl),
         meanCV = mean(cv)) 

save(sim.parallel.seg.perp, file = paste0(Sys.Date(),"PicnicSegPerpspacing",spacing,".Rdata"))

segperp.df <- data.frame(parORseg = "segments",              # parallel grid or segments  
                     perpORparallel = "perpendicular",   # To the gradient
                     spacing = spacing,
                     designAngle = desnAn+90,
                     popsz = N,
                     numSamplers = transects.seg.perp@samp.count,
                     coverageAvg = transects.seg.perp@cov.area/transects.seg.perp@strata.area,
                     summarySegPerp@individuals$summary,
                     summarySegPerp@individuals$N,
                     summarySegPerp@individuals$D,
                     estsegperp.df)

dfout <-  parallel.df %>%
  bind_rows(perp.df) %>%
  bind_rows(seg.df) %>%
  bind_rows(segperp.df)

dfout

}))

save(summarySpacingOrintationType, file = "simSummarySpacing10-25by5.Rdata")


```
